Characteristic polynomials of p-adic matrices

### 2017-02-04 charpoly/charpoly.tex
References
Working on intro, fix typesetting for com, rec, det

\newcommand{\noopsort}[1]{}
\DeclareMathOperator{\NP}{NP}
\DeclareMathOperator{\HP}{HP}
\DeclareMathOperator{\PP}{PP}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\val}{val}
\DeclareMathOperator{\pr}{pr}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator{\com}{Com}
\DeclareMathOperator{\Grass}{Grass}
\DeclareMathOperator{\Lat}{Lat}
\DeclareMathOperator{\round}{round}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\lcm}{lcm}

@@

Instead,

@@

@@

The second ingredient we need is reciprocal polynomials.
We extend its definition to matrices of polynomials.
\begin{deftn}
Let $d \in \mathbb{N}$ and $P \in K[X]$ of degree at most $d.$ 
We define the reciprocal polynomial of order $f$ of $P$ as $P^{rec,d}=X^d P \left( 1/X \right).$
Let $A \in M_n(K[X])$ a matrix of polynomials of degree at most $d.$
We denote by $A^{rec,d}$ the matrix
whose coefficients are the reciprocal polynomials of order $d$ 
of the coefficients of $A.$ 
\end{deftn}
We then have the following result :
\begin{lem}
Let $M \in M_n(K).$ Then:
\begin{eqnarray*}
com(1-XM)^{rec,n-1}=&com(X-M), \\
(\chi_M \times 1)^{rec,n}=&(1-XM) com(1-XM).\\
\end{eqnarray*}
\end{lem}
\begin{proof}
It all comes down to the following result:
let $A \in M_d(K[X])$ a matrix of polynomials of degree at most $1,$
then $det (A^{rec,1})=det(A)^{rec,d}.$
Indeed, one can use multilinearity of the determinant on $X^d det(A(1/X))$
to prove this result.
It directly implies the second part of the lemma.
As for the first part, since coefficients of the comatrix we are interested in
are determinants of a matrix of polynomials of degree at most 1, the result
is also clear.
\end{proof}

This lemma allows us to compute $com(1-XM)$ instead of $com(X-M).$
This has a remarkable advantage: the pivots during the computation of
the SNF of $com(1-XM)$ are units of $O_K[[X]],$ and are known
in advance to be on the diagonal. This leads to a very smooth
precision and complexity behaviour when the entry matrix lives in 
$M_n(O_K).$

@@

\noindent 9. \ \textbf{Return} $V^{rec,n-1}, \psi^{rec,n}.$

@@

\noindent 6. \ Return $com(X-M):=(PT^{-1}S U \times V S^{-1} T P^{-1}) \mod \chi (M).$

>>>>

\newcommand{\noopsort}[1]{}
\DeclareMathOperator{\NP}{NP}
\DeclareMathOperator{\HP}{HP}
\DeclareMathOperator{\PP}{PP}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\val}{val}
\DeclareMathOperator{\pr}{pr}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator{\com}{Com}
\DeclareMathOperator{\Grass}{Grass}
\DeclareMathOperator{\Lat}{Lat}
\DeclareMathOperator{\round}{round}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\lcm}{lcm}
\DeclareMathOperator{\rec}{rec}

@@

Instead, we separate the computation of the precision of $\chi_M$ from the computation
of an approximation to $\chi_M$.  Given some precision on $M$, we use \cite{caruso-roe-vaccon:14a}*{Lem. 3.4}
to find the best possible precision for $\chi_M$.  The analysis of this precision is the subject
of much of this paper.  With this precision known, the actual calculation of $\chi_M$
may proceed by lifting $M$ to a temporarily higher precision and then using a sufficiently
stable algorithm. \todo{reference a place with more detail.}

@@

One benefit of this approach is that we may account for diffuse precision:
precision that is not localized on any single coefficient of $\chi_M$.  For example,
let $0 \le \alpha_1 \le \alpha_2 \le \dots \le \alpha_n$, consider a
diagonal matrix $D$ with diagonal entries $(p^{\alpha_1}, \dots, p^{\alpha_n})$,
let $P, Q \in \GL_n(\Zp)$ and set $M = PDQ$.  The valuation of the coefficient
of $X^{n-k}$ in $\chi_M$ will be $\sum_{i=1}^k \alpha_i$, and if $M$ is known
with precision $O(p^m)$ then $\chi_M$ will be known with some precision
larger than $O(p^m)$ (see Proposition \ref{prop:mainlem}).

@@

The second ingredient we need is reciprocal polynomials.
We extend its definition to matrices of polynomials.
\begin{deftn}
Let $d \in \mathbb{N}$ and $P \in K[X]$ of degree at most $d.$ 
We define the reciprocal polynomial of order $f$ of $P$ as $P^{\rec,d}=X^d P \left( 1/X \right).$
Let $A \in M_n(K[X])$ a matrix of polynomials of degree at most $d.$
We denote by $A^{\rec,d}$ the matrix
whose coefficients are the reciprocal polynomials of order $d$ 
of the coefficients of $A.$ 
\end{deftn}
We then have the following result :
\begin{lem}
Let $M \in M_n(K).$ Then:
\begin{eqnarray*}
\com(1-XM)^{\rec,n-1}=&\com(X-M), \\
(\chi_M \times 1)^{\rec,n}=&(1-XM) \com(1-XM).\\
\end{eqnarray*}
\end{lem}
\begin{proof}
It all comes down to the following result:
let $A \in M_d(K[X])$ a matrix of polynomials of degree at most $1,$
then $\det (A^{\rec,1})=\det(A)^{\rec,d}.$
Indeed, one can use multilinearity of the determinant on $X^d \det(A(1/X))$
to prove this result.
It directly implies the second part of the lemma.
As for the first part, since coefficients of the comatrix we are interested in
are determinants of a matrix of polynomials of degree at most 1, the result
is also clear.
\end{proof}

This lemma allows us to compute $\com(1-XM)$ instead of $\com(X-M).$
This has a remarkable advantage: the pivots during the computation of
the SNF of $\com(1-XM)$ are units of $O_K[[X]],$ and are known
in advance to be on the diagonal. This leads to a very smooth
precision and complexity behaviour when the entry matrix lives in 
$M_n(O_K).$

@@

\noindent 9. \ \textbf{Return} $V^{\rec,n-1}, \psi^{\rec,n}.$

@@

\noindent 6. \ Return $\com(X-M):=(PT^{-1}S U \times V S^{-1} T P^{-1}) \mod \chi (M).$
